\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-03-14 jasa/paper-jasa.tex
Figure -> Fig.

While the optimal decision function \ref{eqn:best-s-clf} is expected to be
learned for the parameter values $\theta_0$ and $\theta_1$ selected in
Algorithm~\ref{alg:training}, it is not clear whether the optimal decision
function can be expected for data generated from  $\theta'_0$ and $\theta'_1$
never jointly encountered during learning. Similarly, it is not clear how the
limited capacity of the classifier may impact the performance the resulting
parameterized decision function. These issues are left as an area for future
work, but preliminary exploration by \cite{Baldi:2016fzo} show that
Algorithm~\ref{alg:training} is an effective practical approach.

\begin{algorithm}[t]
\caption{Learning a parameterized classifier.}\label{alg:training}
\begin{algorithmic}
    \State ${\cal T := \{ \}}$;
    \While{ $\text{size}({\cal L}) < N $ }
        \State Select or draw $\theta_0$, $\theta_1$ from $\Theta_0$, $\Theta_1$;
	    \State Draw $\mathbf{x} \sim p(\mathbf{x}|\theta_0)$;
		\State ${\cal T} := {\cal T} \cup \{ ((\mathbf{x}, \theta_0, \theta_1), y=0) \}$;
		\State Draw $\mathbf{x} \sim p(\mathbf{x}|\theta_1)$;
		\State ${\cal T} := {\cal T} \cup \{ ((\mathbf{x}, \theta_0, \theta_1), y=1) \}$;
    \EndWhile
    \State Learn a single classifier $s(\mathbf{x}; \theta_0, \theta_1)$ from ${\cal T}$.
\end{algorithmic}
\end{algorithm}

@@

As a simple and illustrative example, let us first consider the approximation of
the log-likelihood ratio $\log \left( r(\mathbf{x};\gamma=0.05,\gamma=0) \right)$ between the 1D mixtures
$p(\mathbf{x}|\gamma=0.05)$ and $p(\mathbf{x}|\gamma=0)$ defined as
\begin{align}
p(\mathbf{x}|\gamma) &= (1-\gamma)\frac{p_{c_0}(\mathbf{x}) +  p_{c_1}(\mathbf{x})}{2}   + \gamma \, p_{c_2}(\mathbf{x}),
\end{align}
where $p_{c_0} := {\cal N}(\mu=-2, \sigma^2=0.25^2)$, $p_{c_1} := {\cal N}(\mu=0, \sigma^2=4)$,
$p_{c_2} := {\cal N}(\mu=1, \sigma^2=0.25)$. Samples drawn for the nominal value $\gamma=0.05$ are shown in
Figure~\ref{fig:1a} and used later for inference.

Figure~\ref{fig:p_sij} shows the intermediate stages for the decomposition described in Section~\ref{sec:mixture}.
The blue and green curves show $p_{c'}(\hat{s}_{c,c'}(\mathbf{x}))$,  the distributions for the score for the sub-classifiers for the three pair-wise comparisons of the mixture components. The red curves in Fig.~\ref{fig:p_sij} show the approximation of the density ratio obtained from those distributions.

@@

    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig1c.pdf}
        \caption{$\log \hat r(\mathbf{x})$ using neural network}
        \label{fig:1c}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig1d.pdf}
        \caption{$\log \hat r(\mathbf{x})$ using random forest}
        \label{fig:1d}
    \end{subfigure}

%    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
%      %(or a blank line to force the subfigure onto a new line)
%    \begin{subfigure}[b]{0.3\textwidth}
%        \includegraphics[width=\textwidth]{figures/fig_ratio_dist.png}
%        \caption{Likelihood ratio distributions}
%        \label{fig:decomposed_ratio}
%    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig-ratio-dist.pdf}
        \caption{$p( \log \hat r(\mathbf{x})) \, | \,\gamma = 0.05)$ }
        \label{fig:decomposed_ratio}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig-likelihood.pdf}
        \caption{$-\log \Lambda(\gamma)$ }
        \label{fig:likelihood}
    \end{subfigure}

@@

Figure~\ref{fig:decomposed_ratio} shows the distribution of $ \log \hat r(\mathbf{x}))$ for $\gamma=0.05$
using the decomposed approximation of $\hat r(\mathbf{x}))$ with neural networks and the distribution
of the exact log-likelihood ratio. While there are some artifacts in the distribution in the low-probability regions and the maximum value of the log-likehood ratio is underestimated, the overall shape of the distribution is well approximated.

Finally, we come to the log-likelihood curve 
%shown in Figure~\ref{fig:likelihood}
\begin{equation}
\log \Lambda( \gamma ) =  \log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} \; .
\end{equation}
for the dataset $\mathcal{D}$ shown in  Fig.~\ref{fig:1a}. 
By exploiting the fact that
\begin{equation}
\log \frac{ p({\cal D} | \gamma = 0.05) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} = \log \frac{ p({\cal D} | \gamma = 0.05) }{ p({\cal D} | \gamma = 0 ) } - \log \frac{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma) }{  p({\cal D} | \gamma = 0 ) },
\end{equation}
the generalized likelihood ratio can be computed by evaluating
both terms with respect to a common reference $\gamma=0$
%reusing the decomposed neural network-based approximation from Figure~\ref{fig:1c}.
%In particular, the maximum likelihood estimator (MLE) in the second term
%can be found 
as outlined in Section~\ref{sec:generalized-likelihood-ratio}.
Figure~\ref{fig:likelihood} shows that the exact likelihood curve is very well 
approximated by the method, confirming that
even when the raw classifier does a poor job at modeling the $s^*(\mathbf{x})$, a good approximations of the 
likelihood ratio can still be obtained by calibrating $s(\mathbf{x})$ (and by decomposing
the mixture, if possible).

@@

\bigskip
An advantage of this approach compared to ABC is that the 
classifier and calibration -- computationally intensive 
parts of the approximation  -- are independent of the of the dataset $\mathcal{D}$. 
Thus once trained and calibrated, the approximation can be applied to any dataset $\mathcal{D}$. 
This makes it computationally efficient to perform ensemble tests of the method.

Figure~\ref{fig:2a} shows the empirical distribution of the maximum likelihood estimators (MLEs) from the 
approximate likelihood compared to the distribution of the MLEs from the exact likelihood. 
It clearly demonstrates that in this case the approximate likelihood yields an unbiased estimator with the essentially the same variance as the exact MLE.
%($\mathbb{E} [\hat \gamma] = 0.0501$, as empirically estimated over $1000$
%trials). 
%Note that had the underlying ratio not been accurate enough because of
%poor training or poor calibration, the two distributions would not be exactly
%overlapping, indicating bias. 
In addition to the MLE, we can study the coverage of a confidence interval based on the likelihood ratio test statistic.
This is done by evaluating $-2\log \Lambda(\gamma=0.05)$ for samples drawn from $p(\mathbf{x} | \gamma=0.05)$.
Wilks's theorem states that the distribution of $-2\log \Lambda(\gamma=0.05)$ should follow a $\chi^2_1$ distribution.
%
%From this, the initial $\Lambda$ test statistic can finally
%be evaluated to determine in this case that the null should not be rejected.  
Figure \ref{fig:2b} also confirms this behavior, supporting the applicability of this method for likelihood-based inference techniques in the likelihood-free setting.
% that $-2 \log \Lambda$, as approximated with our method, is asymptotically $\chi^2_1$ distributed,
%as should be the true statistic when the null is true.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5a.pdf}
        \caption{  }
        \label{fig:5a}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5b.pdf}
        \caption{ }
        \label{fig:5b}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5c.pdf}
        \caption{ }
        \label{fig:5c}
    \end{subfigure}
    \caption{Inference from exact and approximate likelihood ratios. The red dot corresponds
    to the true values $(\alpha=1, \beta=-1)$ used to generate $\mathcal{D}$, the green dot is the MLE
    $(\hat \alpha=1.012, \hat \beta=-0.9221)$  from the exact likelihood, while the blue dot is the MLE $(\hat \alpha=1.008, \hat \beta=-1.004)$ from the approximate likelihood.
    1, 2 and 3-$\sigma$ contours are shown in white.
             (\ref{fig:5a}) The exact log-likelihood for the observed data ${\cal D}$.
             (\ref{fig:5b}) The approximate log-likelihood  evaluated on a coarse $15 \times 15$ fixed grid.
             (\ref{fig:5c}) A Gaussian Process surrogate of the approximate log-likelihood ratio estimated from a Bayesian optimization procedure. White dots show the  parameter points sampled during the optimization process.}
    \label{fig:5}
\end{figure}

The observations ${\cal D}$ represented in Figure~\ref{fig:4}
are random samples with $\alpha=1$ and $\beta=-1$. 
Our goal is to infer the values $\alpha$ and $\beta$ based on ${\cal D}$.
We construct the log-likelihood ratio 
\begin{equation}\label{eqn:obj-ex2}
    -2 \log \Lambda(\alpha, \beta ) = -2 \log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }
\end{equation}
that we calculate by exploiting the fact that
\begin{equation}\label{eqn:obj-ex2-expanded}
\log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) } = \log \frac{p({\cal D} | \alpha, \beta)}{ p({\cal D} | \alpha=0, \beta=0) } - \log \frac{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }{ p({\cal D} | \alpha=0, \beta=0)}.
\end{equation}
Following the procedure described in Section~\ref{sec:param-clf}, we 
% to reduce
%computational requirements, both terms in the right hand side of Eqn.~\ref{eqn:obj-ex2-expanded} are evaluated using 
a single 2-layer neural network (with 5+2 inputs and one output node) to form the parameterized classifier $s(\mathbf{x}; \theta_0, \theta_1)$ and fix $\theta_1=(\alpha=0, \beta=0)$.
Since the generative model is not expensive, the classifier output is calibrated
on-the-fly with histograms for  every candidate parameter pair $(\alpha,\beta)$.

Figure~\ref{fig:5a} shows the exact log-likelihood ratio for this dataset, which has an exact MLE at $(\hat \alpha=1.012, \hat \beta=-0.9221)$. Figure~\ref{fig:5b} shows the approximate log-likelihood ratio evaluated on a coarse grid of parameter values. Some roughness in the contours is observed, which is due to primarily to variance introduced in the calibration procedure. In addition to the statistical fluctuations due to finite calibration samples, there are also fluctuations introduced from changes in the binning of the calibration histograms as $\alpha$ and $\beta$ vary. As discussed in Section~\ref{sec:param-calibration}, a parameterized calibration
procedure should ameliorate this issue, but that is left for now as an area for future work.
Nevertheless, optimizing the approximate log-likelihood ratio with
a Bayesian optimization~\citep{brochu2010tutorial} procedure is efficient and effective. 
After $50$ likelihood evaluations, the maximum likelihood estimate is found at $(\hat \alpha=1.008, \hat \beta=-1.004)$. 
While the goal of the Bayesian optimization procedure is to find the maximum likelihood, the posterior mean of the 
internal Gaussian process, shown in Figure~\ref{fig:5c}, is close to the exact log-likelihood ratio illustrated in Figure~\ref{fig:5a}. 
In each case, the true values $\alpha=1$ and $\beta=-1$ are contained within the $1-\sigma$ likelihood contour.
%For comparison, Figure~\ref{fig:5b} shows the approximate log likelihood ratio evaluated on a fixed grid of values. Since $\hat s(\mathbf{x;\alpha,\beta})$ is calibrated on-the-fly,
%roughness in the contours can be observed, in particular due to non-smooth changes in the resulting histograms as $\alpha$ and $\beta$ vary.

>>>>

While the optimal decision function \ref{eqn:best-s-clf} is expected to be
learned for the parameter values $\theta_0$ and $\theta_1$ selected in
Algorithm~\ref{alg:training}, it is not clear whether the optimal decision
function can be expected for data generated from  $\theta'_0$ and $\theta'_1$
never jointly encountered during learning. Similarly, it is not clear how the
limited capacity of the classifier may impact the performance the resulting
parameterized decision function.
Preliminary exploration by \cite{Baldi:2016fzo} shows that
 a uniform grid scan over parameter space is an effective practical approach;
 however, we introduce the distributions $\pi_{\Theta_0}$ and $\pi_{\Theta_1}$
into the Algorithm~\ref{alg:training} to allow for a more
 sophisticated sampling strategy.
% this formulation of Algorithm~\ref{alg:training} allows for a more
% sophisticated sampling strategy.
% These issues are left as an area for future work.
%For this reason, we introduce the distributions $\pi_{\Theta_0}$ and $\pi_{\Theta_1}$
%into the Algorithm~\ref{alg:training} to allow for a more
% sophisticated sampling strategy.

\begin{algorithm}[t]
\caption{Learning a parameterized classifier.}\label{alg:training}
\begin{algorithmic}
    \State ${\cal T := \{ \}}$;
    \While{ $\text{size}({\cal T}) < N $ }
        %\State Select or draw $\theta_0$, $\theta_1$ from $\Theta_0$, $\Theta_1$;
        \State Draw $\theta_0 \sim \pi_{\Theta_0}$;
	    \State Draw $\mathbf{x} \sim p(\mathbf{x}|\theta_0)$;
		\State ${\cal T} := {\cal T} \cup \{ ((\mathbf{x}, \theta_0, \theta_1), y=0) \}$;
        \State Draw $\theta_1 \sim \pi_{\Theta_1}$;
		\State Draw $\mathbf{x} \sim p(\mathbf{x}|\theta_1)$;
		\State ${\cal T} := {\cal T} \cup \{ ((\mathbf{x}, \theta_0, \theta_1), y=1) \}$;
    \EndWhile
    \State Learn a single classifier $s(\mathbf{x}; \theta_0, \theta_1)$ from ${\cal T}$.
\end{algorithmic}
\end{algorithm}

@@

As a simple and illustrative example, let us first consider the approximation of
the log-likelihood ratio $\log \left( r(\mathbf{x};\gamma=0.05,\gamma=0) \right)$ between the 1D mixtures
$p(\mathbf{x}|\gamma=0.05)$ and $p(\mathbf{x}|\gamma=0)$ defined as
\begin{align}
p(\mathbf{x}|\gamma) &= (1-\gamma)\frac{p_{c_0}(\mathbf{x}) +  p_{c_1}(\mathbf{x})}{2}   + \gamma \, p_{c_2}(\mathbf{x}),
\end{align}
where $p_{c_0} := {\cal N}(\mu=-2, \sigma^2=0.25^2)$, $p_{c_1} := {\cal N}(\mu=0, \sigma^2=4)$,
$p_{c_2} := {\cal N}(\mu=1, \sigma^2=0.25)$. Samples drawn for the nominal value $\gamma=0.05$ are shown in
Fig.~\ref{fig:1a} and used later for inference.

Fig.~\ref{fig:p_sij} shows the intermediate stages for the decomposition described in Section~\ref{sec:mixture}.
The blue and green curves show $p_{c'}(\hat{s}_{c,c'}(\mathbf{x}))$,  the distributions for the score for the sub-classifiers for the three pair-wise comparisons of the mixture components. The red curves in Fig.~\ref{fig:p_sij} show the approximation of the density ratio obtained from those distributions.

@@

    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig1c.pdf}
        \caption{$\log \hat r(\hat{s}(\mathbf{x}))$ using neural network}
        \label{fig:1c}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig1d.pdf}
        \caption{$\log \hat r(\hat{s}(\mathbf{x}))$ using random forest}
        \label{fig:1d}
    \end{subfigure}

%    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
%      %(or a blank line to force the subfigure onto a new line)
%    \begin{subfigure}[b]{0.3\textwidth}
%        \includegraphics[width=\textwidth]{figures/fig_ratio_dist.png}
%        \caption{Likelihood ratio distributions}
%        \label{fig:decomposed_ratio}
%    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig-ratio-dist.pdf}
        \caption{$p( \log \hat r(\hat{s}(\mathbf{x}))) \, | \,\gamma = 0.05)$ }
        \label{fig:decomposed_ratio}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{figures/fig-likelihood.pdf}
        \caption{$-\log \Lambda(\gamma)$ }
        \label{fig:likelihood}
    \end{subfigure}

@@

Fig.~\ref{fig:decomposed_ratio} shows the distribution of $ \log \hat r(\mathbf{x}))$ for $\gamma=0.05$
using the decomposed approximation of $\hat r(\mathbf{x}))$ with neural networks and the distribution
of the exact log-likelihood ratio. While there are some artifacts in the distribution in the low-probability regions and the maximum value of the log-likehood ratio is underestimated, the overall shape of the distribution is well approximated.

Finally, we come to the log-likelihood curve
%shown in Figure~\ref{fig:likelihood}
\begin{equation}
\log \Lambda( \gamma ) =  \log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} \;
\end{equation}
for the dataset $\mathcal{D}$ shown in  Fig.~\ref{fig:1a}.
By exploiting the fact that
\begin{equation}
\log \frac{ p({\cal D} | \gamma ) }{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma )} = \log \frac{ p({\cal D} | \gamma ) }{ p({\cal D} | \gamma = 0 ) } - \log \frac{ \sup_{\gamma \in \Theta} p({\cal D} | \gamma) }{  p({\cal D} | \gamma = 0 ) },
\end{equation}
the generalized likelihood ratio can be computed by evaluating
both terms with respect to a common reference $\gamma=0$
%reusing the decomposed neural network-based approximation from Figure~\ref{fig:1c}.
%In particular, the maximum likelihood estimator (MLE) in the second term
%can be found
as outlined in Section~\ref{sec:generalized-likelihood-ratio}.
Fig.~\ref{fig:likelihood} shows that the exact likelihood curve is very well
approximated by the method, confirming that
even when the raw classifier does a poor job at modeling the $s^*(\mathbf{x})$, a good approximations of the
likelihood ratio can still be obtained by calibrating $s(\mathbf{x})$ (and by decomposing
the mixture, if possible).

@@

An advantage of this approach compared to ABC is that the
classifier and calibration -- computationally intensive
parts of the approximation  -- are independent of the of the dataset $\mathcal{D}$.
Thus once trained and calibrated, the approximation can be applied to any dataset $\mathcal{D}$.
This makes it computationally efficient to perform ensemble tests of the method.

Fig.~\ref{fig:2a} shows the empirical distribution of the maximum likelihood estimators (MLEs) from the
approximate likelihood compared to the distribution of the MLEs from the exact likelihood.
It clearly demonstrates that in this case the approximate likelihood yields an unbiased estimator with the essentially the same variance as the exact MLE.
%($\mathbb{E} [\hat \gamma] = 0.0501$, as empirically estimated over $1000$
%trials).
%Note that had the underlying ratio not been accurate enough because of
%poor training or poor calibration, the two distributions would not be exactly
%overlapping, indicating bias.
In addition to the MLE, we can study the coverage of a confidence interval based on the likelihood ratio test statistic.
This is done by evaluating $-2\log \Lambda(\gamma=0.05)$ for samples drawn from $p(\mathbf{x} | \gamma=0.05)$.
Wilks's theorem states that the distribution of $-2\log \Lambda(\gamma=0.05)$ should follow a $\chi^2_1$ distribution.
%
%From this, the initial $\Lambda$ test statistic can finally
%be evaluated to determine in this case that the null should not be rejected.
Fig.~\ref{fig:2b} also confirms this behavior, supporting the applicability of this method for likelihood-based inference techniques in the likelihood-free setting.
% that $-2 \log \Lambda$, as approximated with our method, is asymptotically $\chi^2_1$ distributed,
%as should be the true statistic when the null is true.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5a.pdf}
        \caption{  }
        \label{fig:5a}
    \end{subfigure}
    ~
    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5b.pdf}
        \caption{ }
        \label{fig:5b}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig5c.pdf}
        \caption{ }
        \label{fig:5c}
    \end{subfigure}
    \caption{Inference from exact and approximate likelihood ratios. The red dot corresponds
    to the true values $(\alpha=1, \beta=-1)$ used to generate $\mathcal{D}$, the green dot is the MLE
    $(\hat \alpha=1.012, \hat \beta=-0.9221)$  from the exact likelihood, while the blue dot is the MLE $(\hat \alpha=1.008, \hat \beta=-1.004)$ from the approximate likelihood.
    1, 2 and 3-$\sigma$ contours are shown in white.
             (\ref{fig:5a}) The exact log-likelihood for the observed data ${\cal D}$.
             (\ref{fig:5b}) The approximate log-likelihood  evaluated on a coarse $15 \times 15$ fixed grid.
             (\ref{fig:5c}) A Gaussian Process surrogate of the approximate log-likelihood ratio estimated from a Bayesian optimization procedure. White dots show the  parameter points sampled during the optimization process.}
    \label{fig:5}
\end{figure}

The observations ${\cal D}$ represented in Fig.~\ref{fig:4}
are random samples with $\alpha=1$ and $\beta=-1$.
Our goal is to infer the values $\alpha$ and $\beta$ based on ${\cal D}$.
We construct the log-likelihood ratio
\begin{equation}\label{eqn:obj-ex2}
    -2 \log \Lambda(\alpha, \beta ) = -2 \log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }
\end{equation}
that we calculate by exploiting the fact that
\begin{equation}\label{eqn:obj-ex2-expanded}
\log \frac{p({\cal D} | \alpha, \beta)}{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) } = \log \frac{p({\cal D} | \alpha, \beta)}{ p({\cal D} | \alpha=0, \beta=0) } - \log \frac{ \sup_{\alpha,\beta} p({\cal D} | \alpha, \beta) }{ p({\cal D} | \alpha=0, \beta=0)}.
\end{equation}
Following the procedure described in Section~\ref{sec:param-clf}, we build
% to reduce
%computational requirements, both terms in the right hand side of Eqn.~\ref{eqn:obj-ex2-expanded} are evaluated using
a build single 2-layer neural network (with 5+2 inputs and one output node) to form the parameterized classifier $s(\mathbf{x}; \theta_0, \theta_1)$ and fix $\theta_1=(\alpha=0, \beta=0)$.
Since the generative model is not expensive, the classifier output is calibrated
on-the-fly with histograms for  every candidate parameter pair $(\alpha,\beta)$.

Figure~\ref{fig:5a} shows the exact log-likelihood ratio for this dataset, which has an exact MLE at $(\hat \alpha=1.012, \hat \beta=-0.9221)$.
Figure~\ref{fig:5b} shows the approximate log-likelihood ratio evaluated on a coarse grid of parameter values. Some roughness in the contours is observed, which is due to primarily to variance introduced in the calibration procedure. In addition to the statistical fluctuations due to finite calibration samples, there are also fluctuations introduced from changes in the binning of the calibration histograms as $\alpha$ and $\beta$ vary. As discussed in Section~\ref{sec:param-calibration}, a parameterized calibration
procedure should ameliorate this issue, but that is left for now as an area for future work.
Nevertheless, optimizing the approximate log-likelihood ratio with
a Bayesian optimization~\citep{brochu2010tutorial} procedure is efficient and effective.
After $50$ likelihood evaluations, the maximum likelihood estimate is found at $(\hat \alpha=1.008, \hat \beta=-1.004)$.
While the goal of the Bayesian optimization procedure is to find the maximum likelihood, the posterior mean of the
internal Gaussian process, shown in Fig.~\ref{fig:5c}, is close to the exact log-likelihood ratio illustrated in Fig.~\ref{fig:5a}.
In each case, the true values $\alpha=1$ and $\beta=-1$ are contained within the $1-\sigma$ likelihood contour.
%For comparison, Figure~\ref{fig:5b} shows the approximate log likelihood ratio evaluated on a fixed grid of values. Since $\hat s(\mathbf{x;\alpha,\beta})$ is calibrated on-the-fly,
%roughness in the contours can be observed, in particular due to non-smooth changes in the resulting histograms as $\alpha$ and $\beta$ vary.
